### 2024-05-13 paper/draft-introduction.tex
updated introduction

\documentclass{article}
\usepackage{graphicx} % Required for inserting images
\usepackage[margin=0.75in]{geometry}
\usepackage{outlines}

\begin{document}

\section{Introduction}

Inland water bodies present a unique challenge to characterization by remote sensing imagery due to their complex spectral characteristics and small scale spatial variability. The broad bands of multi-spectral imagers coupled with the irregular shape of lakes and rivers results in pixels with highly mixed signals that are easily dominated by reflectance from shore and near-shore vegetation sources \cite{koponen2002lake, ritchie2003remote}. Recently, the combination of hyperspectral imaging with unmanned aerial vehicles (UAV) such as drones has emerged as a powerful approach to simultaneously address the spectral, spatial, and temporal limitations of traditional high-altitude and satellite-based collection \cite{adao2017hyperspectral,arroyo2019implementation}. UAV are significantly less expensive to deploy than their remote sensing counterparts, and low altitude flights enable centimeter-scale sampling while limiting the need for complicated atmospheric corrections \cite{banerjee2020uav}. However, the significant increase in data volume generated by these systems presents a new challenge, namely, how to efficiently extract water quality parameters of interest from intricate pixel spectra.

Significant research efforts have focused on the development of techniques and algorithms to retrieve water quality parameters from UAV-captured hyperspectral images (HSI). On-board compute installed alongside hyperspectral imagers can enable the rapid evaluation of spectral indices from HSI band ratios \cite{horstrand2019uav}. These band ratios and polynomial combinations of bands have been used to successfully invert optically-active water quality parameters such as turbidity directly from UAV acquired imagery \cite{vogt2016near, zhang2022selection}. Supervised machine learning techniques such as tree-based models, support vector machines, and neural networks have also been used to estimate a plethora of parameters such as colored dissolved organic matter, chlorophyll a, blue-green algae, and suspended sediment concentrations \cite{keller2018hyperspectral, lu2021retrieval}. The calibration and evaluation of these data-driven models demand a significant volume of coincident, in situ data. This can be addressed by coordinating UAV flights with reference data collection using autonomous robotic boats \cite{robot-team-1, robot-team-2}. Nevertheless, this approach relies on prior knowledge of expected sources in order to select appropriate reference instruments for model validation. The presence of unanticipated contaminants cannot be directly identified in this sensing paradigm.

Extending the capabilities of UAV-based hyperspectral imaging to enable water quality monitoring in real-time scenarios where contaminant sources may not be known in advance requires two additional capabilities: dimensionality reduction techniques to permit the visual comparison of HSI, and endmember extraction techniques which can identify spectral signatures corresponding to unique sources within the imaging scene. In remote sensing where reference data are typically sparse, many approaches have been explored. For example, principal component analysis (PCA) and t-distributed stochastic neighbor embedding (tSNE) are dimensionality reduction methods commonly used to reduce HSI to two or three dimensions for visualization \cite{tyo2003principal,zhang2015hyperspectral}. Similarly for endmember extraction there are a variety of established approaches including geometric methods like vertex component analysis, statistical methods like k-nearest neighbors and non-negative matrix factorization (NMF), and deep learning methods based on autoencoder architectures \cite{heylen2014review,nascimento2005vertex, Feng2022HyperspectralUB, cariou2015unsupervised, su2019daen, borsoi2019deep, palsson2020convolutional}. Methods based on linear mappings like PCA and NMF are often too restrictive for HSI where the assumption of linear mixing is easily broken. However, the increased complexity of nonlinear methods like tSNE and autoencoders often lead to significant increases in computation time. An ideal approach should enable both visualization and nonlinear identification of relevant spectral endmembers.

% neurons that wire together, fire together
The self-organizing map (SOM) developed by Teuvo Kohonen is an unsupervised machine learning method which nonlinearly maps high dimensional data to the nodes of a two-dimensional grid \cite{kohonen-som-1}. By preserving the topological relationship between nodes during training, the SOM ensures that similar spectral signatures are mapped together such that related HSI pixels naturally cluster together. This presents an attractive compromise by enabling the simultaneous visualization of HSI data and endmember extraction via the weight vector associated with each SOM node \cite{cantero2004analysis, duran2007time,som-hsi}. When reference data are available, the SOM can be utilized to provide semi-supervised labeling of HSI spectra \cite{riese2019supervised}. Furthermore, Danielsen et al. demonstrated that the dimensionality reduction offered by the SOM can even be used for on-board data compression of HSIs acquired by a CubeSat \cite{danielsen2021self}. Despite these clear capabilities, the SOM relies on a heuristic training algorithm with hyperparemeters that can be challenging to tune and offers no direct probabilistic interpretation. To address these limitations, Bishop et al. developed the generative topographic mapping (GTM), a Bayesian latent-variable model inspired by the SOM \cite{gtm-bishop-1}. The GTM has been utilized in a variety of domains including drug design and chemical data visualization but has yet to see adoption for the analysis of hyperspectral imagery \cite{kireeva2012generative, gaspar2015chemical, horvath2019generative}.

In this paper, we explore the application of the GTM to UAV-acquired HSI for the characterization water quality. Specifically, we use the GTM to

Using data collected at site in Montague, North Texas, we demonstrate the ability of this paradigm with two case studies. First, we demonstrate how the GTM can be used to provide a high-resolution unsupervised classification of a water body enabling the rapid identification of relevant regions which can then be used to direct further investigation by reference-sensor equipped boats. Second, we demonstrate the ability to perform endmember extraction and abundance mapping using imagery from a test-release of rhodamine dye. Together with the normalized spectral similarity score, these spectral signatures are identified and their associated area in the scene is determined.

\bibliographystyle{plain}
\bibliography{paper/references.bib}

\end{document}